---
title: "COVID-19 Canadian data: pre-processing, visualizations, and modelling"
author: "Stephen Pierzchajlo"
output:
  #rmdformats::readthedown:
  #rmdformats::robobook: 
  #html_document:
  github_document:
  #rmdformats::html_clean:
  #rmdformats::html_docco:
  #prettydoc::html_pretty:
    #theme: cayman
    #theme: tactile
    #theme: architect
    #theme: leonids
    #theme: hpstr
  #rmdformats::material: 
    #self_contained: true
    #thumbnails: false
    #lightbox: true
    #gallery: false
    #highlight: tango
    #code_folding: hide
---

# Introduction

I have used R for a few years now, and every project I am currently working on has a completed R script already made. Therefore, it didn't make sense to me to just redo my analyses. I decided I wanted to do something new for this course, and do something I had never done before. One aspect of programming that I never was comfortable with was writing functions. So the ultimate goal of this project is to write functions that will perform an entire analysis in an interesting way, while still staying within the parameters of the assignment.

## My project idea

Many months ago I analysed data for a paper on COVID-19. The paper is currently under review, but all analyses, data, and scripts can be found here: https://osf.io/c8zhn/. Part of my work involved getting COVID-19 infection rate data for 4 specific weeks in March/April to use as a random effects parameter in a multiple regression model. The specific COVID-19 repository I found gets updated daily and contains data for every country (and every sub-region in that country) in the world. I thought it would be interesting to create a set of functions that could analyse any country the user chooses in a variety of ways, and output COVID-19 related statistics and visualisations of infection trends. So that is the ultimate goal of this project.

<br/>

In this project, I will first analyse COVID-19 data for 1 country in great detail, and then ultimately automate these analyses with a number of functions that generalise across any country the user wants. Because I am Canadian, I decided to first analyse Canadian COVID-19 data in as much detail as possible. This will include downloading the data from a specific source, wrangling it into a condensed dataframe, graphing the data in various ways, and even doing some basic modelling to predict future infection rates.

# Preparations

Here is how I typically start any R project.

## Remove global environment

I typically remove the global environment everytime I work on a project.

```{r remove global environment, message=FALSE, warning=FALSE}

# This function removes the entire global environment!!!
rm(list = ls())

```

<br/>

## Load libraries

Below is a list of libraries I will use. I comment on each one to indicate what they will be used for.

```{r load libraries, message=FALSE, warning=FALSE}

# Load libraries.

# Contains all tidyverse function for data wrangling and graphing.
library(tidyverse)

# Used for graphing daily COVID-19 infections.
library(gghighlight)

# Summarising data into condensed dataframe.
library(plyr)

# Theme for ggplot that I like.
#devtools::install_github('Mikata-Project/ggthemr')
library(ggthemr)

# THe ggarrange function plots multiple graphs.
library(ggpubr)

# Display coefficients for predictive models.
library(broom)

# Create documentation for functions.
library(docstring)

```

<br/>

# Data wrangling

## Finding a COVID-19 repository

Their are many different COVID-19 data repositories, and I am using two files that contains COVID-19 data for every country in the world. File 1 contains data about total worldwide cases and file 2 contains data about total worldwide deaths. Each time the code chunk below is run, it grabs each *.csv* file from a github repository that has been updated within the past 24 hours. This is ideal, because it means that anytime I run this script, I have COVID-19 statistics that are as recent as yesterday. I prefer this approach to downloading a static *.csv* file because COVID-19 cases and deaths are ongoing and constantly changing.

```{r load data, message=FALSE, warning=FALSE}

# Load .csv files from github. These are updated daily, so everytime this analysis is run, it uses the most up to date information.

# Load total global cases data.
COVID_Global_Wide <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

# Load total global deaths data.
COVID_Global_Deaths_Wide <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))

```

<br/>

## Quick look at raw data

First I'll look at the overall shape of one of these dataframes, since I already know they both have the same structure.

```{r print rows/columns, message=FALSE, warning=FALSE}

# Print number of rows and columns.
cat(" There are", nrow(COVID_Global_Wide), "rows in this dataframe.\n","There are", ncol(COVID_Global_Wide), "columns in this dataframe.")

```

There are a massive number of columns. Why is that? I decided to take a look at the overall structure of one of the dataframes.

```{r quick look, message=FALSE, warning=FALSE}

# Display first 10 rows and all columns.
COVID_Global_Wide[1:10, ]

```

As seen above, the *.csv* file has been arranged such that each column is a date (ignoring the first 4 columns), with values in a column corresponding to the total cases/deaths up until that date. This format is not easy to work with, and it means a new column will appear every day. After 1 week this dataframe will have 7 more columns.

<br/>

## Reshaping dataframes

To make the data easier to work with, I will create one column called "date" where each row is a date. In a column beside it, I will have the total cases/deaths for that date. Now instead of 100s of columns that contain dates, there is instead a single dates column.

```{r reshape raw data, message=FALSE, warning=FALSE}

# Total global cases data in long format.
COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)

# Total global deaths data in long format
COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)

```

```{r Inspect, message=FALSE, warning=FALSE}

# Take a look at several of the first rows.
head(COVID_Global_Long_Date)

# Take a look at several of the first rows.
head(COVID_Global_Deaths_Long_Date)

```

Now both dataframes contain a single dates column.

<br/>

## Combining dataframes

Since both dataframes come from the same source, they might have the same structure. If so, I could just attach the deaths column from the second dataframe to the first dataframe. Below I loop through both dataframe, and for each column check how many rows have matching information in the same position, and how many have non-matching information. I do this for all pairs of columns except for the cases/deaths (since those should be different). If there are any differences, I should get some false statments below. If there are no differences, I should have a single count that says "true".

```{r check similarity, message=FALSE, warning=FALSE}

# Check whether the cases and deaths dataframes have exactly the same information per row (except actual cases/deaths).
for(i in length(colnames(COVID_Global_Long_Date)) - 2){
  
  print(table(COVID_Global_Long_Date[[i]] == COVID_Global_Deaths_Long_Date[[i]]))
  
}

```

Everything matches, so I'm going to add the cases column from the deaths dataframe to the cases dataframe. Additionally, I'll call this new column "deaths":

```{r combine dataframes, message=FALSE, warning=FALSE}

# Combine death count to cases dataframe.
COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases

```

Now, I can quickly check the dataframe to see how it looks.

```{r quick check 2, message=FALSE, warning=FALSE}

# See first few rows.
head(COVID_Global_Long_Date)

```

Now I only need to work with this single dataframe instead of 2.

<br/>

## Filtering provinces

Because I only want Covid-19 data from Canada, I need to filter out every other non-Canadian country.

```{r subset 1, message=FALSE, warning=FALSE}

# Subset Canadian provinces from total cumulative global cases data.
Covid_Canada <- subset(COVID_Global_Long_Date, `Country/Region` == "Canada")

```

I can then check how many provinces I have in the filtered dataframe.

```{r check provinces 1, message=FALSE, warning=FALSE}

# See how many provinces are in the dataset.
unique(Covid_Canada$`Province/State`)

```

<br/>

Now there are only 16 provinces in this dataframe, covering every Canadian province. However, I only want to focus on the most populated provinces (many territories have extremely small populations). Therefore, I will focus on the 8 largest Canadian provinces, which will be subsetted below.

```{r subset 2, message=FALSE, warning=FALSE}

# List of Provinces.
Province_List <- c("Ontario", "Quebec", "New Brunswick", "Manitoba", "British Columbia", "Saskatchewan", "Alberta", "Newfoundland and Labrador")

# For province data, I'm not immediately interested in Canadian Territories, so I'll filter those out.
Covid_Canada <- Covid_Canada %>%
  filter(`Province/State` %in% Province_List)

```

Now I can double-check the number of provinces in my further filtered dataframe.

```{r check provinces 2, message=FALSE, warning=FALSE}

# See how many provinces are in the dataset.
unique(Covid_Canada$`Province/State`)

```

Looks good!

<br/>

I also want to filter out dates earlier than March 1st 2020. 

```{r filter march, message=FALSE, warning=FALSE}

# Filter canadian provincial data to only include data starting from March 1st.
Covid_Canada <- Covid_Canada[352:nrow(Covid_Canada), ]

# Check first row of date column,
Covid_Canada[1,5]

```

It appears now that the dataframe starts around the first of march. My filtering wasn't exact, but I think it's good enough.

<br/>

## Converting dates

Looking at the data below...

```{r look at date, message=FALSE, warning=FALSE}

# Check data structure.
str(Covid_Canada)

```

...the date object is a Factor. However, I want it to be a Date object instead.

```{r change to date, message=FALSE, warning=FALSE}

# Convert date column to date format for cases dataframe.
Covid_Canada$date <- as.Date(Covid_Canada$date, format = "%m/%d/%Y")

# Display date variable type.
str(Covid_Canada$date)

```

Now we the dates are formatted as actual Date objects.

<br/>

## Subsetting dataframe

I also don't need the Country/Region, Lat, or Long columns, so I'm going to get rid of them.

```{r remove columns, message=FALSE, warning=FALSE}

# Take only a subset of columns.
Covid_Canada <- subset(Covid_Canada, select = c("Province/State", "date", "cases", "deaths"))

```

<br/>

## Daily cases/deaths calculation

The Covid-19 cases and deaths data are cumulative, so each row represents the total cases or deaths there have been up to that date. If I want cases per day, I need to add new columns that subtract the current day from the previous day. Below I make two new columns that only have cases/deaths for that day. To do that I take a value from one row, and subtract it from the previous row.

```{r daily cases/deaths, message=FALSE, warning=FALSE}

# Calculate cases per day by subtracting current count in each row by total cases up until that point.
Covid_Canada$DailyCases <- ave(Covid_Canada$cases,Covid_Canada$`Province/State`,  FUN=function(x) c(0, diff(x)))

# Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
Covid_Canada$DailyDeaths <- ave(Covid_Canada$deaths,Covid_Canada$`Province/State`,  FUN=function(x) c(0, diff(x)))

```

<br/>

## Per Capita Correction

Because I am comparing provinces, absolute cases/deaths will be misleading since different provinces have different population sizes. I am going to add a new column using case_when() that gives cases/deaths per 100,000 people. I had to look up the population for each province, and then I divided cases/deaths in each province by the province's population, and multiplied it by 100,000.

```{r per capita cases, message=FALSE, warning=FALSE}

# Calculate total cases in each province per 100,000 people.
Covid_Canada$Cases_Per100000 <- as.numeric(case_when(Covid_Canada$`Province/State` == "Alberta" ~ (Covid_Canada$cases/4345737) * 100000,
                                                     Covid_Canada$`Province/State` == "British Columbia" ~ (Covid_Canada$cases/5020302) * 100000,
                                                     Covid_Canada$`Province/State` == "Manitoba" ~ (Covid_Canada$cases/1360396) * 100000,
                                                     Covid_Canada$`Province/State` == "New Brunswick" ~ (Covid_Canada$cases/772094) * 100000,
                                                     Covid_Canada$`Province/State` == "Newfoundland and Labrador" ~ (Covid_Canada$cases/523790) * 100000,
                                                     Covid_Canada$`Province/State` == "Ontario" ~ (Covid_Canada$cases/14446515) * 100000,
                                                     Covid_Canada$`Province/State` == "Quebec" ~ (Covid_Canada$cases/8433301) * 100000,
                                                     Covid_Canada$`Province/State` == "Saskatchewan" ~ (Covid_Canada$cases/1168423) * 100000))

```

```{r per capita deaths, message=FALSE, warning=FALSE}

# Calculate total deaths in each province per 100,000 people.
Covid_Canada$Deaths_Per100000 <- as.numeric(case_when(Covid_Canada$`Province/State` == "Alberta" ~ (Covid_Canada$deaths/4345737) * 100000,
                                                      Covid_Canada$`Province/State` == "British Columbia" ~ (Covid_Canada$deaths/5020302) * 100000, 
                                                      Covid_Canada$`Province/State` == "Manitoba" ~ (Covid_Canada$deaths/1360396) * 100000,
                                                      Covid_Canada$`Province/State` == "New Brunswick" ~ (Covid_Canada$deaths/772094) * 100000,
                                                      Covid_Canada$`Province/State` == "Newfoundland and Labrador" ~ (Covid_Canada$deaths/523790) * 100000,
                                                      Covid_Canada$`Province/State` == "Ontario" ~ (Covid_Canada$deaths/14446515) * 100000,
                                                      Covid_Canada$`Province/State` == "Quebec" ~ (Covid_Canada$deaths/8433301) * 100000,
                                                      Covid_Canada$`Province/State` == "Saskatchewan" ~ (Covid_Canada$deaths/1168423) * 100000))

```

<br/>

# Graphs

## Total cases/deaths in each province.

I want to first see the total cases/deaths in each province. I decided to make a stacked bar graph where each bar represents total cases in a province. Each bar is seperated by the proportion of cases to deaths. The x-axis displays total cases, and I put text beside each bar to display just the total deaths per province.

```{r bar graph, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# average data in dataframe for graphing.
Covid_Canada %>%
  ddply(c("`Province/State`"), summarise,
        cases = sum(DailyCases),
        deaths = sum(DailyDeaths)) %>%
  gather(Category, Number, cases:deaths) %>%
  
  # Graph total cases and deaths per province, tilt sideways, stack cases/deaths, display cases count on x-axis/death count above each bar.
  ggplot(aes(x = reorder(`Province/State`, Number), y = Number, fill = factor(Category, levels = c("deaths", "cases")))) +
  geom_bar(stat = "identity") + 
  scale_colour_ggthemr_d() +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))  +
  labs(x = "Province",
       y = "COVID-19 Cases And Deaths",
       title = "Total COVID-19 Cases in Canada",
       caption = c("Data source: Government of Canada", paste0("Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1))),
       fill = "Case Type") +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = ifelse(Category=="deaths", Number,"")), position = position_stack(vjust=1.1), hjust = -.2, color = "blue") +
  coord_flip()

```

<br/>

## Cumulative cases/deaths in each province {.tabset}

Below I have a visualisation of the cumulative cases/deaths for each province.

### Cumulative cases

```{r per capita cases graph, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# Cumulative Cases Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = Cases_Per100000, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_area() +
  labs(x = paste0(head(format(Covid_Canada$date, "%B %d %y"), n = 1)," to Present"),
       y = "Cumulative Cases Per 100,000 People",
       title = "Coronavirus Cases Per Canadian Province",
       caption = c("Data source: Government of Canada", paste0("Today's Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1)))) +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  scale_colour_ggthemr_d() +
  scale_x_date(date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))

```

<br/>

### Cumulative deaths

Below I have a visualisation of the cumulative deaths for each province.

```{r per capita deaths graph, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

#Cumulative Deaths Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = Deaths_Per100000, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_area() +
  labs(x = paste0(head(format(Covid_Canada$date, "%B %d %y"), n = 1)," to Present"),
       y = "Cumulative Deaths Per 100,000 People",
       title = "Coronavirus Deaths Per Canadian Province",
       caption = c("Data source: Government of Canada", paste0("Today's Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1)))) +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  scale_colour_ggthemr_d() +
  scale_x_date(date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))

```

<br/>

## Daily cases/deaths per province {.tabset}

Below are daily cases and deaths for each province. I used gghighlight to superimpose total cases/deaths behind each provinces cases/deaths.

### Daily cases

```{r daily cases graph, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# Daily Cases Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = DailyCases, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_bar(stat = "identity") +
  facet_wrap(~`Province/State`, nrow = 2, ncol = 4) +
  labs(x = paste0(head(format(Covid_Canada$date, "%B %d %y"), n = 1)," to Present"),
       y = "Cases Per Day",
       title = "Daily Coronavirus Cases Per Canadian Province",
       caption = c("Data source: Government of Canada", paste0("Today's Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1)))) +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  scale_colour_ggthemr_d() +
  scale_x_date(date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5)) +
  gghighlight()

```

<br/>

### Daily Deaths

```{r daily deaths graph, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# Daily Deaths Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = DailyDeaths, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_bar(stat = "identity") +
  facet_wrap(~`Province/State`, nrow = 2, ncol = 4) +
  labs(x = paste0(head(format(Covid_Canada$date, "%B %d %y"), n = 1)," to Present"),
       y = "Deaths Per Day",
       title = "Daily Coronavirus Deaths Per Canadian Province",
       caption = c("Data source: Government of Canada", paste0("Today's Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1)))) +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  scale_colour_ggthemr_d() +
  scale_x_date(date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5)) +
  gghighlight()

# Reset theme.
ggthemr_reset()

```

<br/>

# Modelling
Because the assignment calls for some sort of analysis, I decided to try some predictive modelling using this dataset. While this dataset lends itself much better to qualitative analysis, I have seen many COVID-19 models that try to predict future cases based on current trends. I have never done anything like that before, but I thought this might be a good way to learn. 

<br/>

## Data Preparation

First, I made a new dataframe for predictive modelling, since I will need to add future dates to it.

```{r message=FALSE, warning=FALSE}

# Make modelling dataframe.
Covid_Canada_Modelling <- Covid_Canada

```

<br/>

I also want to make a prediction model for all of Canada, so I am summarising by date so that each day averages over all cases/deaths for every province.

```{r message=FALSE, warning=FALSE}

# Average cases per day, for all provinces.
Covid_Canada_dates <- Covid_Canada_Modelling %>%
  ddply(c("date"), summarise,
        cases = mean(cases),
        deaths = mean(deaths),
        DailyCases = sum(DailyCases),
        DailyDeaths = sum(DailyDeaths))

```

<br/>

## Create future predictions dataframe

Most researchers who try to predict future COVID-19 cases use more complicated models with many parameters to account for things like lockdowns, mask compliance, etc. I don't have any predictors like that in the current dataset, so it will be hard to predict anything too far into the future. However, even with a single potentially mediocre predictor like date, I may be able to get an accurate estimate of very short-term trends. Therefore, I will try to predict COVID-19 cases for the next 10 days. To do this, I'm going to make a new dataframe that has the next 10 days after the most recent date today. I will also had blank columns that match the main dataframe so that I can bind their rows later.

```{r message=FALSE, warning=FALSE}

# New dataframe that starts the final day of COVID data collection and adds 20 days. Also contains blank columns for other variables.
New_Dates <- data.frame(date =seq(max(Covid_Canada_Modelling$date),max(Covid_Canada_Modelling$date) + 10,by='days'),
                        cases = NA,
                        deaths = NA,
                        DailyCases = NA,
                        DailyDeaths = NA)

# View future dates dataframe.
head(New_Dates)

```

The dataframe just contains a column of future dates, and NA's for all other columns. One issue is that this dataframe starts on the last day that COVID-19 data has been collected, since I created the dates by starting with the last date on the main dataframe. So I will just delete the first row from this new dataframe.

```{r message=FALSE, warning=FALSE}

# Delete first row.
New_Dates = New_Dates[-1,]

# Look at first few rows.
head(New_Dates)

```

Now the dataframe starts on the NEXT day.

<br/>

Next, I just add this dataframe to the modelling dataframe.

```{r message=FALSE, warning=FALSE}

# Add future dates to end of modelling dataframe.
Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)

# Look at last 15 rows of ammended dataframe.
tail(Covid_Canada_dates, n = 15)

```

Now the dataframe contains future dates as well.

<br/>

## Data modelling

Next I need to define the actual model. Unfortunately, I don't have any useful predictors other than date, and just using a date to predict the number of COVID-19 cases is probably severely lacking. Still, using current infection trends from the past might be a good way to estimate very near future infection rates (like infections 5-10 days from now). Therefore, I will set up a model that only uses date to predict infections for the next 10 days.

<br/>

### Naive model

First, I will fit a very naive model. This model uses a single predictor (dates), and uses data from the past 300 days of Covid-19 infections.

```{r message=FALSE, warning=FALSE}

# Somewhat naive linear model.
model.1 <- lm(data = tail(Covid_Canada_dates, n = 300), cases ~ date)

```

<br/>

I can then look at a summary of the coefficients I estmated for my model, and print out an interpretation.

```{r message=FALSE, warning=FALSE}

# Model summary.
tidy(model.1, conf.int = T)

# Print coefficient interpretation.
print(paste0("For every 1 day increase over the past 300 days, we would expect an additional ", round(tidy(model.1, conf.int = T)[2,2], 2), " new Covid-19 infections"))

```

<br/>

### Prediction

I now have a simple model, so I will apply those coefficients to the new days. The model is this: 

```{r message=FALSE, warning=FALSE}

# Print predictive model summary with coefficients.
print(paste0("cases = ", round(tidy(model.1, conf.int = T)[1,2], 2), " + ", round(tidy(model.1, conf.int = T)[2,2], 2), " * date"))

```

Thus, this model can be used to predict new cases, by simply inputing the current date into this equation and calculating cases.

```{r message=FALSE, warning=FALSE}

# Predict COVID-19 infections with my model.
Covid_Canada_dates$New <-  predict(object = model.1, newdata = Covid_Canada_dates)

# Look at predictions for future dates.
tail(Covid_Canada_dates$New, n = 10)

```

Each value above represents predictions from my model for the next 10 days. I can quickly see how close my predictions are to the actual COVID-19 cases for those dates, as well as see what the model's future predicitons are by looking at the final 10 of the dataframe.

```{r message=FALSE, warning=FALSE}

# Look at predictions for future dates.
data.frame("Date" = tail(Covid_Canada_dates$date, n = 15),
           "Actual Cases" = tail(Covid_Canada_dates$cases, n = 15),
           "Predicted Cases" = tail(Covid_Canada_dates$New, n = 15))

```

The predicted cases don't match the actual cases too well. I'll plot the regression line over the data to see what the disparity might be.

```{r message=FALSE, warning=FALSE}

# Plot data and future predictions regression lne.
ggplot(data = tail(Covid_Canada_dates, n = 300), aes(x = date, y = cases)) + 
  geom_point(color='blue')  +
  geom_line(color='red',data = tail(Covid_Canada_dates, n = 300), aes(x=date, y=New))

```

The regression line (red) extends beyond the data (blue), so it's definitely predicting future infections. But, the model isn't a very good fit as the straight line doesn't fit the non-linear data very well. It also seems to be underpredicting Covid-19 infections from January on. This is because a single-degree regression model with just one predictor doesn't allow for any flexibility if the data are even slightly non-linear. So, I will instead fit a series of polynomial models that can increasingly account for the non-linearity seen in the data.

<br/>

## Polynomial models 1

In the code chunk below, I will repeat the steps I performed above, but this time fit a series of polynomial models. This will be done in a for loop, from 1 degree to 8 degrees. Finally, I will plot each model against the data.

```{r message=FALSE, warning=FALSE}

# Make 8 polynomial models, graph them, and collect adjusted R-Square value for each model.
for(i in 1:8){
  
  # Make modelling dataframe. 
  Covid_Canada_dates <- Covid_Canada
  
  # Average cases per day, for all provinces.
  Covid_Canada_dates <- Covid_Canada_dates %>%
    ddply(c("date"), summarise, 
          cases = mean(cases), 
          deaths = mean(deaths),
          DailyCases = sum(DailyCases),
          DailyDeaths = sum(DailyDeaths))
  
  # New dates dataframe.
  New_Dates <- data.frame(date =seq(max(Covid_Canada_dates$date),max(Covid_Canada_dates$date) + 30,by='days'),
                          cases = NA,
                          deaths = NA,
                          DailyCases = NA,
                          DailyDeaths = NA)
  
  # Delete first row.
  New_Dates = New_Dates[-1,]
  
  # Add future dates to end of modelling dataframe.
  Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)
  
  # Fit increasingly comlex polynomials.
  model.1 <- lm(data = tail(Covid_Canada_dates, n = 300), cases ~ poly(date, i))

  # Predict COVID-19 infections with each polynomial model.
  Covid_Canada_dates$New <-  predict(object = model.1, newdata = Covid_Canada_dates)
  
  # Assign new name to each graph I make below.
  assign(paste0("Graph", i),
         
         # Plot data and polynomial regression line.
         ggplot(data = tail(Covid_Canada_dates, n = 300), aes(x = date, y = cases)) + 
           geom_point(color='blue')  +
           ggtitle(paste0(i, " degree polynomial")) +
           theme(plot.title = element_text(size=10)) +
           theme(plot.title = element_text(hjust = 0.5)) +
           geom_line(color='red',data = tail(Covid_Canada_dates, n = 300), aes(x=date, y=New)))
  
  # Get Adjusted R-Squared value for each polynomial and assign it a new name.
  assign(paste0("Adj", i), summary(model.1)$adj.r.squared)
  
  # Model Summary.
  assign(paste0("Model_Coeffs", i),broom::tidy(model.1, conf.int = T))
  
}

```

Now I should be able to plot all 8 graphs in a single image.

```{r message=FALSE, warning=FALSE, fig.height=6, fig.width=7}

# Plot all 8 models on a single graph.
multi <- ggarrange(Graph1, Graph2, Graph3, Graph4, Graph5, Graph6, Graph7, Graph8)

# Annotate multiple graphs.
annotate_figure(multi,top = text_grob("Predicting Next 10 Days Of COVID", color = "black", face = "bold", size = 14))

```

Each model fits differently, with some models fitting the data better than others, and some models making what appear to be more reasonable future predictions.. However, these models may also be severly overfitting to the data, which one would expect as more parameters are added to a model. In addition, I decided that using dates from the past 8 months might not be best when trying to assess short term trends in infection rates. For instance, infections from 8 months ago have far less bearing on future trends than do infections from 2 weeks ago. Therefore, I am going to instead fit these polynomials to only a small number of days in the past.

<br/>

## Polynomial models 2

Here I will repeat the loop above, but this time predict the next 10 days of infection using only the previous 30 days as a predictor.

```{r message=FALSE, warning=FALSE}

# Make 8 polynomial models, graph them, and collect adjusted R-Square value for each model.
for(i in 1:8){
  
  # Make modelling dataframe.
  Covid_Canada_dates <- Covid_Canada
  
  # Average cases per day, for all provinces.
  Covid_Canada_dates <- Covid_Canada_dates %>%
    ddply(c("date"), summarise,
          cases = mean(cases),
          deaths = mean(deaths),
          DailyCases = sum(DailyCases),
          DailyDeaths = sum(DailyDeaths))
  
  # New dates dataframe.
  New_Dates <- data.frame(date =seq(max(Covid_Canada_dates$date),max(Covid_Canada_dates$date) + 10,by='days'),
                          cases = NA,
                          deaths = NA,
                          DailyCases = NA,
                          DailyDeaths = NA)
  
  # Delete first row.
  New_Dates = New_Dates[-1,]

  # Add future dates to end of modelling dataframe.
  Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)

  # Fit increasingly comlex polynomials.
  model.2 <- lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, i))
  
  assign(paste0("model_", i), lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, i)))

  #broom::tidy(model.2, conf.int = T)

  # Predict COVID-19 infections with each polynomial model.
  Covid_Canada_dates$New <-  predict(object = model.2, newdata = Covid_Canada_dates)
  
  # Assign new name to each graph I make below.
  assign(paste0("Graph", i),
         
         # Plot data and polynomial regression line.
         ggplot(data = tail(Covid_Canada_dates, n = 30), aes(x = date, y = cases)) + 
           geom_point(color='blue')  +
           ggtitle(paste0(i, " degree polynomial")) +
           theme(plot.title = element_text(size=10)) +
            theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1), 
                  plot.title = element_text(hjust = 0.5)) +
           geom_line(color='red',data = tail(Covid_Canada_dates, n = 30), aes(x=date, y=New)))
  
  # Get Adjusted R-Squared value for each polynomial and assign it a new name.
  assign(paste0("Adj", i), summary(model.2)$adj.r.squared)
  
  # 
  assign(paste0("Mod", i), lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, i)))
  
  # Model Summary.
  assign(paste0("Model_Coeffs", i),broom::tidy(model.1, conf.int = T))

  }


```

Again, I can now plot all 8 graphs as 1 image.

```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=6}

# Plot all 8 models on a single graph.
multi <- ggarrange(Graph1, Graph2, Graph3, Graph4, Graph5, Graph6, Graph7, Graph8)

# Annotate multiple graphs.
annotate_figure(multi,top = text_grob("Predicting Next 10 Days Of COVID", color = "black", face = "bold", size = 14))

```

Some of the higher order polynomial fits seem to be making absurdly unbelievable future predictions. However, some of the lower-order polynomial models look reasonable. I'm going to focus on the 3-order polynomial (which looks alright) and pretend it was the model I thought was best.

```{r message=FALSE, warning=FALSE}

# Make modelling dataframe.
Covid_Canada_Modelling <- Covid_Canada

# Average cases per day, for all provinces.
Covid_Canada_dates <- Covid_Canada_Modelling %>%
  ddply(c("date"), summarise,
        cases = mean(cases),
        deaths = mean(deaths),
        DailyCases = sum(DailyCases),
        DailyDeaths = sum(DailyDeaths))

# New dataframe that starts the final day of COVID data collection and adds 20 days. Also contains blank columns for other variables.
New_Dates <- data.frame(date =seq(max(Covid_Canada_Modelling$date),max(Covid_Canada_Modelling$date) + 10,by='days'),
                        cases = NA,
                        deaths = NA,
                        DailyCases = NA,
                        DailyDeaths = NA)

# Delete first row.
New_Dates = New_Dates[-1,]

# Add future dates to end of modelling dataframe.
Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)

model.2 <- lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, 3))

# Model summary.
tidy(model.2, conf.int = T)

# Predict COVID-19 infections with my model.
Covid_Canada_dates$New <-  predict(object = model.1, newdata = Covid_Canada_dates)

```

I could then apply these coefficients to each new date and get a new COVID-19 infection total for that day in the future.

```{r message=FALSE, warning=FALSE}

# Look at predictions for future dates.
data.frame("Date" = tail(Covid_Canada_dates$date, n = 15),
           "Actual Cases" = tail(Covid_Canada_dates$cases, n = 15),
           "Predicted Cases" = tail(Covid_Canada_dates$New, n = 15))

```

The predicted data matches the actual cases pretty closely for days that have case data. Thus, the future predictions might now be a little bit more accurate.

<br/>

It should be stressed, however, that this model still probably isn't very good. A more accurate model would need many more predictors, and I would probably assess overfitting using a process like cross-validation to compute model variance out of sample. But for this assignment, I will leave this as it is.

<br/>

# Functions  {.tabset}

The dataset I used contains data from every country in the world, not just Canada. If I wanted to repeat the graphs and analyses I did, I would have to painstakingly replace each code chunk with the new country and re-run everything. I decided to make some functions that not only repeat the analyses I did for Canada, but allow the user to simply add the country they want to the function, and instantly get up to date COVID graphs and numbers for that country.

<br/>

## Country List

For users of these functions, it will be useful to get a list of names of countries that can be used. This function takes no arguments, and simply outputs a list of all countries the user could choose from.

```{r message=FALSE, warning=FALSE}

Country_List <- function(){
  
  #'@title Country_List: finding countries for COVID-19 plotting in R.
  #'
  #'@author Stephen Pierzchajlo
  #'
  #'@description Function that returns a list of countries
  #' 
  #'@usage Country_List()
  #'
  #'@details This function contains no arguments. The user simply types the function as 
  #'outlined above under Usage, and recieves a list of all countries in the world that can 
  #'be used with other subsequent functions.
  
  # Make package list.
  packages <- c('tidyverse')    
  
  # Install packages not yet installed
  installed_packages <- packages %in% rownames(installed.packages())
  
  if (any(installed_packages == FALSE)) {
    
    install.packages(packages[!installed_packages], repos = c(CRAN = "https://cran.rstudio.com"))
    
  }
  
  # Packages loading
  invisible(lapply(packages, library, character.only = TRUE))
  
  # Total Global Cases Data
  x <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(x, date, cases, 5:ncol(x), factor_key=TRUE)
  
  # compute all unique countries and return a list.
  return(unique(COVID_Global_Long_Date$`Country/Region`))
  
}

```


## Total cases/deaths in total country

This function outputs the number of cases and deaths up until the current date in the country chosen by the user. It is only a text output, and probably is not that important since the next function will output a graph, but I thought it was nice to include in case someone wants to view it.

```{r message=FALSE, warning=FALSE}

COVID_Country_Data <- function(country){
  
  #'@title COVID_Country_Data: raw COVID-19 cases and deaths.
  #'
  #'@author Stephen Pierzchajlo
  #'
  #'@description Function that returns raw COVID-19 cases and deaths for a specific country.
  #' 
  #'@usage COVID_Country_Data(country)
  #'
  #'@param country character. Name of country
  #'
  #'@details This function is used to get a numerical output of cases and deaths for a specific 
  #'country. Use Country_List() for a list of country names that can be passed to this function.
  
  # Make package list.
  packages <- c('tidyverse', 'plyr')    
  
  # Install packages not yet installed
  installed_packages <- packages %in% rownames(installed.packages())
  
  if (any(installed_packages == FALSE)) {
    
    install.packages(packages[!installed_packages], repos = c(CRAN = "https://cran.rstudio.com"))
    
  }
  
  # Packages loading
  invisible(lapply(packages, library, character.only = TRUE))
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
  
  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  ### Some countries have provinces, others don't. The strategy I am using to make the dataframes for entire country
  ### won't work in both instances. So we need to apply averaging differently.
  
  # If Province/State is empty:
  if(is.na(Covid$`Province/State`) == TRUE){
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
  }
  
  # If Province/State has values:
  else{
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
  }
  
  # Make total cases/deaths chart.
  Chart <- Covid %>%
    ddply(c("`Country/Region`"), summarise,
          cases = sum(DailyCases),
          deaths = sum(DailyDeaths)) %>%
    gather(Category, Number, cases:deaths)
  
  # Return chart.
  return(Chart)
  
}

```


## COVID Country Graph Function

```{r message=FALSE, warning=FALSE}

COVID_Country_Graph <- function(country){
  
  #'@title COVID_Country_Graph: Plot total COVID-19 cases and deaths.
  #'
  #'@author Stephen Pierzchajlo
  #'
  #'@description Function that plots the total COVID-19 cases and deaths in a specific country.
  #' 
  #'@usage COVID_Country_Graph(country)
  #'
  #'@param country character. Name of country
  #'
  #'@details This function returns a single graph containing total COVID-19 cases 
  #'and deaths per day. Use Country_List() for a list of country names that can 
  #'be passed to this function.
  
  # Make package list.
  packages <- c('tidyverse', 'plyr', 'ggthemr')    
  
  # Install packages not yet installed
  installed_packages <- packages %in% rownames(installed.packages())
  
  if (any(installed_packages == FALSE)) {
    
    install.packages(packages[!installed_packages], repos = c(CRAN = "https://cran.rstudio.com"))
    
  }
  
  # Packages loading
  invisible(lapply(packages, library, character.only = TRUE))
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
  
  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  ### Some countries have provinces, others don't. The strategy I am using to make the dataframes for entire country
  ### won't work in both instances. So we need to apply averaging differently.
  
  # If Province/State is empty:
  if(is.na(Covid$`Province/State`) == TRUE){
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
  }
  
  # If Province/State has values:
  else{
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
  }
  
  # Set Theme.
  ggthemr("flat")
  
  # average data in dataframe for graphing.
  a <- Covid %>%
    ddply(c("`Country/Region`"), summarise,
          cases = sum(DailyCases),
          deaths = sum(DailyDeaths)) %>%
    gather(Category, Number, cases:deaths) %>%
    
    # Graph total cases and deaths per province, tilt sideways, stack cases/deaths, display cases count on x-axis/death count above each bar.
    ggplot(aes(x = `Country/Region`, y = Number, fill = factor(Category, levels = c("deaths", "cases")))) +
    geom_bar(stat = "identity", position = "dodge") + 
    scale_colour_ggthemr_d() +
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE))  +
    labs(x = "Country",
         y = "COVID-19 Cases And Deaths",
         title = paste0("Total COVID-19 Cases In ", country),
         caption = paste0("Data source: Government of ", country),
         fill = "Case Type") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label = Number), position = position_dodge(width=0.9), vjust = -.2)
  
  return(a)
  
}

```


## COVID Country Cumulative Graph Function

```{r message=FALSE, warning=FALSE}

COVID_Country_Cumulative_Graph <- function(country){
  
  #'@title COVID_Country_Cumulative_Graph: cumulative COVID-19 cases and deaths.
  #'
  #'@author Stephen Pierzchajlo
  #'
  #'@description Function that plots cumulative COVID-19 cases and deaths in a specific country.
  #' 
  #'@usage COVID_Country_Cumulative_Graph(country)
  #'
  #'@param country character. Name of country
  #'
  #'@details This function returns 2 graphs, each containing either cumulative COVID-19 cases 
  #'or cumulative COVID-19 deaths. Use Country_List() for a list of country names that can 
  #'be passed to this function.
  
  # Make package list.
  packages <- c('tidyverse', 'plyr', 'ggthemr')    
  
  # Install packages not yet installed
  installed_packages <- packages %in% rownames(installed.packages())
  
  if (any(installed_packages == FALSE)) {
    
    install.packages(packages[!installed_packages], repos = c(CRAN = "https://cran.rstudio.com"))
    
  }
  
  # Packages loading
  invisible(lapply(packages, library, character.only = TRUE))
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
  
  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Filter canadian provincial data to only include data starting from March 1st.
  Covid <- COVID_Global_Long_Date[352:nrow(COVID_Global_Long_Date), ]
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  # Convert date column to date format for cases dataframe.
  Covid$date <- as.Date(Covid$date, format = "%m/%d/%Y")
  
  # Calculate cases per day by subtracting current count in each row by total cases up until that point.
  Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
  
  # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
  Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
  
  # Set Theme.
  ggthemr("flat")
  
  # Cumulative Cases Graph.
  cases <- Covid %>% 
    ggplot(aes(x = date, y = cases, color = `Country/Region`, group = `Country/Region`, fill = `Country/Region`)) +
    geom_area() +
    labs(x = paste0(head(format(Covid$date, "%B %d %y"), n = 1)," to Present"),
         y = "Cumulative Cases",
         title = paste0("Cumulative Coronavirus Cases In ", country),
         caption = c(paste0("Data source: Government of ", country), paste0("Today's Date: ", tail(format(Covid$date, "%B %d %y"), n = 1)))) +
    scale_colour_ggthemr_d() +
    scale_x_date(date_breaks = '1 month', date_labels = '%B') +
    theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))
  
  deaths <- Covid %>% 
    ggplot(aes(x = date, y = deaths, color = `Country/Region`, group = `Country/Region`, fill = `Country/Region`)) +
    geom_area() +
    labs(x = paste0(head(format(Covid$date, "%B %d %y"), n = 1)," to Present"),
         y = "Cumulative Cases",
         title = paste0("Cumulative Coronavirus Deaths In ", country),
         caption = c(paste0("Data source: Government of ", country), paste0("Today's Date: ", tail(format(Covid$date, "%B %d %y"), n = 1)))) +
    scale_colour_ggthemr_d() +
    scale_x_date(date_breaks = '1 month', date_labels = '%B') +
    theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))
  
  output <- list(cases, deaths)
  
  return(output)
  
}

```


## COVID Country Daily Graph Function

```{r message=FALSE, warning=FALSE}

COVID_Country_Daily_Graph <- function(country){
  
  #'@title COVID_Country_Daily_Graph: plot daily COVID-19 cases and deaths.
  #'
  #'@author Stephen Pierzchajlo
  #'
  #'@description Function that plots daily COVID-19 cases and deaths in a specific country.
  #' 
  #'@usage COVID_Country_Daily_Graph(country)
  #'
  #'@param country character. Name of country
  #'
  #'@details This function returns 2 graphs, each containing either COVID-19 cases per day 
  #'or COVID-19 deaths per day. Use Country_List() for a list of country names that can 
  #'be passed to this function.
  
  # Make package list.
  packages <- c('tidyverse', 'plyr', 'ggthemr')    
  
  # Install packages not yet installed
  installed_packages <- packages %in% rownames(installed.packages())
  
  if (any(installed_packages == FALSE)) {
    
    install.packages(packages[!installed_packages], repos = c(CRAN = "https://cran.rstudio.com"))
    
  }
  
  # Packages loading
  invisible(lapply(packages, library, character.only = TRUE))
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
  
  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Filter canadian provincial data to only include data starting from March 1st.
  Covid <- COVID_Global_Long_Date[352:nrow(COVID_Global_Long_Date), ]
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  # Convert date column to date format for cases dataframe.
  Covid$date <- as.Date(Covid$date, format = "%m/%d/%Y")
  
  ### Some countries have provinces, others don't. The strategy I am using to make the dataframes for entire country
  ### won't work in both instances. So we need to apply averaging differently.
  
  # If Province/State is empty:
  if(is.na(Covid$`Province/State`) == TRUE){
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    Covid <- Covid %>%
      ddply(c("`date`"), summarise,
            cases = sum(DailyCases),
            deaths = sum(DailyDeaths))
    
  }
  
  # If Province/State has values:
  else{
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    Covid <- Covid %>%
      ddply(c("`date`"), summarise,
            cases = sum(DailyCases),
            deaths = sum(DailyDeaths))
    
  }
  
  # Set Theme.
  ggthemr("flat")
  
  # Daily Cases Graph.
  cases <- ggplot(Covid, aes(x = date, y = cases)) +
    geom_bar(stat = "identity") +
    labs(x = paste0(head(format(Covid$date, "%B %d %y"), n = 1)," to Present"),
         y = "Cases Per Day",
         title = paste0("Daily Coronavirus Cases In ", country),
         caption = c(paste0("Data source: Government of ", country), paste0("Today's Date: ", tail(format(Covid$date, "%B %d %y"), n = 1)))) +
    theme(plot.caption = element_text(hjust=c(0, 1))) +
    scale_colour_ggthemr_d() +
    scale_x_date(date_breaks = '1 month', date_labels = '%B') +
    theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))
  
  # Daliy Deaths Graph.
  deaths <- ggplot(Covid, aes(x = date, y = deaths)) +
    geom_bar(stat = "identity") +
    labs(x = paste0(head(format(Covid$date, "%B %d %y"), n = 1)," to Present"),
         y = "Deaths Per Day",
         title = paste0("Daily Coronavirus Deaths In ", country),
         caption = c(paste0("Data source: Government of ", country), paste0("Today's Date: ", tail(format(Covid$date, "%B %d %y"), n = 1)))) +
    theme(plot.caption = element_text(hjust=c(0, 1))) +
    scale_colour_ggthemr_d() +
    scale_x_date(date_breaks = '1 month', date_labels = '%B') +
    theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))
  
  outputs <- list(cases, deaths)
  
  return(outputs)
  
}

```

# Testing Functions

Here I am going to use each function as if I had not done any of the previous work in this R project. So, I will first delete everything that has been done up to this point.

```{r remove global environment 2nd time, message=FALSE, warning=FALSE}

# This function removes the entire global environment!!!
rm(list = ls())

```

Next, I will load the docstring package manually since it cannot be loaded from inside a function. The docstring package allows you to add documentation to a function. If it isn't installed, install before running the next line.

```{r load docstring, message=FALSE, warning=FALSE}
# Create documentation for functions.
library(docstring)
```

## Source Functions

I made 5 functions, so I will link to their directories so they can be automatically called.

```{r message=FALSE, warning=FALSE}
source("C:\\Users\\STPI0560\\Desktop\\R Projects\\COVID-19---Canada\\bin\\Country_List.R", local = knitr::knit_global())

source("C:\\Users\\STPI0560\\Desktop\\R Projects\\COVID-19---Canada\\bin\\COVID_Country_Data.R", local = knitr::knit_global())

source("C:\\Users\\STPI0560\\Desktop\\R Projects\\COVID-19---Canada\\bin\\COVID_Country_Graph.R", local = knitr::knit_global())

source("C:\\Users\\STPI0560\\Desktop\\R Projects\\COVID-19---Canada\\bin\\COVID_Country_Cumulative_Graph.R", local = knitr::knit_global())

source("C:\\Users\\STPI0560\\Desktop\\R Projects\\COVID-19---Canada\\bin\\COVID_Country_Daily_Graph.R", local = knitr::knit_global())
```

## Trying Functions Out

Here I will try the functions out with a few countries other than canada to see what the output looks like. First, I will look a the list of countries and pick a few.

```{r message=FALSE, warning=FALSE}

Country_List()

```

From the list above, I will pick a few countries. Israel has seen cases plummet, so I would like to see their trends. Brazil has seen an exponential rise in cases, so I would like to see this as well.

```{r message=FALSE, warning=FALSE}

COVID_Country_Data(Israel)
COVID_Country_Data(Brazil)


```


```{r message=FALSE, warning=FALSE}

COVID_Country_Graph(Israel)
COVID_Country_Graph(Brazil)

```


```{r message=FALSE, warning=FALSE}

COVID_Country_Cumulative_Graph(Israel)
COVID_Country_Cumulative_Graph(Brazil)

```


```{r message=FALSE, warning=FALSE}

COVID_Country_Daily_Graph(Israel)
COVID_Country_Daily_Graph(Brazil)

```