---
title: "COVID-19 Canadian data: pre-processing, visualizations, and modelling"
author: "Stephen Pierzchajlo"
output: github_document
  #html_document:
  #github_document:
    
#always_allow_html: true

---

# Introduction

Many months ago I helped analyse data for a paper on COVID-19. The paper is currently under review, but all analyses, data, and scripts can be found here: https://osf.io/c8zhn/

<br/>

My contribution included getting COVID-19 data for 4 specific weeks to use as 1 predictor in a model for the main analysis of the paper. Because I have a COVID-19 data repository at hand, I wanted to put together a project where I look at Canadian COVID-19 data.

<br/>

Here I'll demonstrate how to automatically download COVID-19 worldwide data, clean the dataframe, display simple figures, and do some simple predictive modelling. Importantly, the data I'll be using is massive and not yet ready for visualisation in its raw format, so I need to perform some wrangling first.

```{r remove global environment, message=FALSE, warning=FALSE, echo = FALSE}

# This function removes the entire global environment!!!
rm(list = ls())

```

## Load Libraries

```{r load libraries, message=FALSE, warning=FALSE}

# Load libraries
library(tidyverse)
library(gghighlight)
library(plyr)
library(ggthemr)
library(ggpubr)
library(broom)

```

<br/>

## Finding a COVID-19 Repository

Their are different COVID-19 data repositories, and I am using two files that contains COVID-19 data for every country in the world. File 1 contains data about total worldwide cases and file 2 contains data about total worldwide deaths. Each time the code chunk below is run, it grabs each *.csv* file from a github repository that has been updated within the past 24 hours. This is ideal, because it means that anytime I run this script, I have COVID-19 statistics that are as recent as yesterday. I prefer this approach to downloading a static *.csv* file because COVID-19 cases and deaths are ongoing and constantly changing.

```{r load data, message=FALSE, warning=FALSE}

# Here, we load .csv files from github. These are updated daily, so everytime this analysis is run, it uses the most up to date information.

# Total Global Cases Data
COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

# Total Global Deaths Data
COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))

```

<br/>

## Quick Look at Raw Data
First I'll look at the overall shape of one of these dataframes.

```{r print rows/columns, message=FALSE, warning=FALSE}

# Print number of rows and columns.
cat(" There are", nrow(COVID_Global_Wide), "rows in this dataframe.\n","There are", ncol(COVID_Global_Wide), "columns in this dataframe.")

```

There are a massive number of columns. Why is that? I decided to take a look at the overall structure.

```{r quick look, message=FALSE, warning=FALSE}

# Display first 10 rows and all columns.
COVID_Global_Wide[1:10, ]

```

As seen above, the *.csv* file has been arranged such that each column is a date (ignoring the first 4 columns), with values in a column corresponding to the total cases/deaths up until that date. This format is not easy to work with, and it means a new column will appear every day. After 1 week this dataframe will have 7 more columns.

<br/>

## Reshaping Dataframes
To make the data easier to work with, I will create one column called "date" where each row is a date. In a column beside it, I will have the total cases/deaths for that date. Now instead of 100s of columns that contain dates, we only have one column.

```{r reshape raw data, message=FALSE, warning=FALSE}

# Total global cases data in long format.
COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)

# Total global deaths data in long format
COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)

# Take a look at several of the first rows.
head(COVID_Global_Long_Date)

```

<br/>

## Combining Dataframes
Since both dataframes come from the same source, they might have the same structure. If so, we could just attach the deaths column from the second dataframe to the first dataframe. Below we loop through both dataframe, and for each column check how many rows have matching information in the same position, and how many have non-matching information. We do this for all pairs of columns except for the cases/deaths (since those should be different). If there are any differences, we should get some false statments below. If there are no differences, we should have a single count that says "true".

```{r check similarity, message=FALSE, warning=FALSE}

# Check whether the cases and deaths dataframes have exactly the same information per row (except actual cases/deaths).
for(i in length(colnames(COVID_Global_Long_Date)) - 2){
  
  print(table(COVID_Global_Long_Date[[i]] == COVID_Global_Deaths_Long_Date[[i]]))
  
}

```

Everything matches, so I'm going to add the cases column from the deaths dataframe to the cases dataframe. Additionally, I'll call this new column "deaths":

```{r combine dataframes, message=FALSE, warning=FALSE}

# Combine death count to cases dataframe.
COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases

```

Now, I can quickly check the dataframe to see how it looks.

```{r quick check 2, message=FALSE, warning=FALSE}

# See first few rows.
head(COVID_Global_Long_Date)

```

<br/>

## Filtering Provinces
Because I only want data from Canada, I need to filter out every other country.

```{r subset 1, message=FALSE, warning=FALSE}

# Subset Canadian provinces from total cumulative global cases data.
Covid_Canada <- subset(COVID_Global_Long_Date, `Country/Region` == "Canada")

```

I can then check how many provinces I have in the filtered dataframe.

```{r check provinces 1, message=FALSE, warning=FALSE}

# See how many provinces are in the dataset.
unique(Covid_Canada$`Province/State`)

```

<br/>

Now, their are only 16 provinces, covering every Canadian province. However, I only want to focus on the most populated provinces (many territories have extremely small populations). Therefore, I will focus on the 8 largest provinces, which will be subsetted below.

```{r subset 2, message=FALSE, warning=FALSE}

# List of Provinces.
Province_List <- c("Ontario", "Quebec", "New Brunswick", "Manitoba", "British Columbia", "Saskatchewan", "Alberta", "Newfoundland and Labrador")

# For province data, I'm not immediately interested in Canadian Territories, so I'll filter those out.
Covid_Canada <- Covid_Canada %>%
  filter(`Province/State` %in% Province_List)

```

Now I can double-check the number of provinces in my further filtered dataframe.

```{r check provinces 2, message=FALSE, warning=FALSE}

# See how many provinces are in the dataset.
unique(Covid_Canada$`Province/State`)

```

Looks good!

<br/>

I also want to filter out dates earlier than March 1st 2020. 

```{r filter march, message=FALSE, warning=FALSE}

# Filter canadian provincial data to only include data starting from March 1st.
Covid_Canada <- Covid_Canada[352:nrow(Covid_Canada), ]

# Check first row of date column,
Covid_Canada[1,5]

```

It appears now that the dataframe starts around the first of march. My filtering wasn't exact, but I think it's good enough.

<br/>

## Converting Date Variable to Date Object.

Looking at the data below...

```{r look at date, message=FALSE, warning=FALSE}

# Check data structure.
str(Covid_Canada)

```

...the date object is a Factor. However, I want it to be a date object instead. Below I do just that.

```{r change to date, message=FALSE, warning=FALSE}

# Convert date column to date format for cases dataframe.
Covid_Canada$date <- as.Date(Covid_Canada$date, format = "%m/%d/%Y")

# Display date variable type.
str(Covid_Canada$date)

```

Now we can see how the dates are formatted as actual dates.

<br/>

I also don't need the Country/Region, Lat, or Long data, so I'm going to get rid of it.

```{r remove columns, message=FALSE, warning=FALSE}

# Take only a subset of columns.
Covid_Canada <- subset(Covid_Canada, select = c("Province/State", "date", "cases", "deaths"))

```

<br/>

The cases and deaths are cumulative, so each row tells us how many total cases there have been up to that date. If I want cases per day, I need to add new columns. Below I make two new columns that only have cases/deaths for that day. To do that I take a value from one row, and subtract it from the previous row.

```{r daily cases/deaths, message=FALSE, warning=FALSE}

# Calculate cases per day by subtracting current count in each row by total cases up until that point.
Covid_Canada$DailyCases <- ave(Covid_Canada$cases,Covid_Canada$`Province/State`,  FUN=function(x) c(0, diff(x)))

# Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
Covid_Canada$DailyDeaths <- ave(Covid_Canada$deaths,Covid_Canada$`Province/State`,  FUN=function(x) c(0, diff(x)))

```

<br/>

## Per Capita Correction

Because I am comparing provinces, absolute cases/deaths will be misleading since different provinces have different population sizes. I am going to add a new column using case_when, that gives cases/deaths per 100,000 people. I had to look up the population for each province, and then I divided cases/deaths in each province by the province's population, and multiplied it by 100,000.
```{r per capita cases, message=FALSE, warning=FALSE}

# Calculate total cases in each province per 100,000 people.
Covid_Canada$Cases_Per100000 <- as.numeric(case_when(Covid_Canada$`Province/State` == "Alberta" ~ (Covid_Canada$cases/4345737) * 100000,
                                                     Covid_Canada$`Province/State` == "British Columbia" ~ (Covid_Canada$cases/5020302) * 100000,
                                                     Covid_Canada$`Province/State` == "Manitoba" ~ (Covid_Canada$cases/1360396) * 100000,
                                                     Covid_Canada$`Province/State` == "New Brunswick" ~ (Covid_Canada$cases/772094) * 100000,
                                                     Covid_Canada$`Province/State` == "Newfoundland and Labrador" ~ (Covid_Canada$cases/523790) * 100000,
                                                     Covid_Canada$`Province/State` == "Ontario" ~ (Covid_Canada$cases/14446515) * 100000,
                                                     Covid_Canada$`Province/State` == "Quebec" ~ (Covid_Canada$cases/8433301) * 100000,
                                                     Covid_Canada$`Province/State` == "Saskatchewan" ~ (Covid_Canada$cases/1168423) * 100000))

```

```{r per capita deaths, message=FALSE, warning=FALSE}

# Calculate total deaths in each province per 100,000 people.
Covid_Canada$Deaths_Per100000 <- as.numeric(case_when(Covid_Canada$`Province/State` == "Alberta" ~ (Covid_Canada$deaths/4345737) * 100000,
                                                      Covid_Canada$`Province/State` == "British Columbia" ~ (Covid_Canada$deaths/5020302) * 100000, 
                                                      Covid_Canada$`Province/State` == "Manitoba" ~ (Covid_Canada$deaths/1360396) * 100000,
                                                      Covid_Canada$`Province/State` == "New Brunswick" ~ (Covid_Canada$deaths/772094) * 100000,
                                                      Covid_Canada$`Province/State` == "Newfoundland and Labrador" ~ (Covid_Canada$deaths/523790) * 100000,
                                                      Covid_Canada$`Province/State` == "Ontario" ~ (Covid_Canada$deaths/14446515) * 100000,
                                                      Covid_Canada$`Province/State` == "Quebec" ~ (Covid_Canada$deaths/8433301) * 100000,
                                                      Covid_Canada$`Province/State` == "Saskatchewan" ~ (Covid_Canada$deaths/1168423) * 100000))

```

<br/>

# Graphs

## Total cases/deaths in each province.
I want to first see the total cases/deaths in each province. I decided to make a stacked bar graph where each bar represents total cases in a province. Each bar is seperated by the proportion of cases to deaths. The x-axis displays total cases, and I put text beside each bar to display just the total deaths per province.

```{r bar graph, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# average data in dataframe for graphing.
Covid_Canada %>%
  ddply(c("`Province/State`"), summarise,
        cases = sum(DailyCases),
        deaths = sum(DailyDeaths)) %>%
  gather(Category, Number, cases:deaths) %>%
  
  # Graph total cases and deaths per province, tilt sideways, stack cases/deaths, display cases count on x-axis/death count above each bar.
  ggplot(aes(x = reorder(`Province/State`, Number), y = Number, fill = factor(Category, levels = c("deaths", "cases")))) +
  geom_bar(stat = "identity") + 
  scale_colour_ggthemr_d() +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))  +
  labs(x = "Province",
       y = "COVID-19 Cases And Deaths",
       title = "Total COVID-19 Cases in Canada",
       caption = c("Data source: Government of Canada", paste0("Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1))),
       fill = "Case Type") +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = ifelse(Category=="deaths", Number,"")), position = position_stack(vjust=1.1), hjust = -.2, color = "blue") +
  coord_flip()

```

```{r per capita cases graph, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")
# Cumulative Cases Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = Cases_Per100000, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_area() +
  labs(x = paste0(head(format(Covid_Canada$date, "%B %d %y"), n = 1)," to Present"),
       y = "Cumulative Cases Per 100,000 People",
       title = "Coronavirus Cases Per Canadian Province",
       caption = c("Data source: Government of Canada", paste0("Today's Date: ", tail(format(Covid_Canada$date, "%B %d %y"), n = 1)))) +
  theme(plot.caption = element_text(hjust=c(0, 1))) +
  scale_colour_ggthemr_d() +
  scale_x_date(date_breaks = '1 month',
        date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

```

```{r per capita deaths graph, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

#Cumulative Deaths Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = Deaths_Per100000, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_area() +
  labs(x = "March 1st (2020) To Present",
       y = "Cumulative Deaths Per 100,000 People",
       title = "Coronavirus Deaths Per Canadian Province",
       caption = "Data source: Government of Canada") +
  scale_colour_ggthemr_d() +
  scale_x_date(date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

```

```{r daily cases graph, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# Daily Cases Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = DailyCases, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_bar(stat = "identity") +
  facet_wrap(~`Province/State`, nrow = 2, ncol = 4) +
  labs(y = "Cases Per Day",
       title = "Daily Coronavirus Cases Per Canadian Province",
       caption = "Data source: Government of Canada") +
  scale_colour_ggthemr_d() +
  scale_x_date(name = 'March 1st (2020) To Present', date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  gghighlight()

```


```{r daily deaths graph, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, dpi=300}

# Set Theme.
ggthemr("flat")

# Daily Deaths Graph.
Covid_Canada %>% 
  ggplot(aes(x = date, y = DailyDeaths, color = `Province/State`, group = `Province/State`, fill = `Province/State`)) +
  geom_bar(stat = "identity") +
  facet_wrap(~`Province/State`, nrow = 2, ncol = 4) +
  labs(y = "Deaths Per Day",
       title = "Daily Coronavirus Deaths Per Canadian Province",
       caption = "Data source: Government of Canada") +
  scale_colour_ggthemr_d() +
  scale_x_date(name = 'March 1st (2020) To Present', date_breaks = '1 month', date_labels = '%B') +
  theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  gghighlight()

# Reset Theme.
ggthemr_reset()

```

# Modelling
Because the assignment calls for some sort of analysis, I decided to try some predictive modelling using this dataset. While this dataset lends itself much more to qualitative analysis, I have seen many COVID models where they try to predict future cases based on current trends. I have never done anything like that before, but I thought this might be a good way to learn. 

<br/>

First, I Made a new dataframe for predictive modelling, since I will need to add future dates to it.
```{r message=FALSE, warning=FALSE}

# Make modelling dataframe.
Covid_Canada_Modelling <- Covid_Canada

```

<br/>

I also want to make a prediction model for all of Canada, so I am summarising by date so that each day averages over all cases/deaths for every province.

```{r message=FALSE, warning=FALSE}

# Average cases per day, for all provinces.
Covid_Canada_dates <- Covid_Canada_Modelling %>%
  ddply(c("date"), summarise,
        cases = mean(cases),
        deaths = mean(deaths),
        DailyCases = sum(DailyCases),
        DailyDeaths = sum(DailyDeaths))

```

<br/>

Most models that try to predict future COVID-19 cases use more complicated models with many parameters to account for things like lockdowns, mask compliance, etc. I don't have anything like that here, so I am only going to try and predict cases for the next 10 days. Therefore I need to add the next 20 days to the dates column, and leave all other columns blank.

```{r message=FALSE, warning=FALSE}

# New dataframe that starts the final day of COVID data collection and adds 20 days. Also contains blank columns for other variables.
New_Dates <- data.frame(date =seq(max(Covid_Canada_Modelling$date),max(Covid_Canada_Modelling$date) + 10,by='days'),
                        cases = NA,
                        deaths = NA,
                        DailyCases = NA,
                        DailyDeaths = NA)

# View future dates dataframe.
New_Dates

```

We can see the dataframe just contains a column of future dates, and NA's for all other columns. This is so that we can add these dates to the modelling dataframe we already have. However, this dataframe also starts on the last day that we have COVID-19 data for, since I created the dates by starting with the last date on that dataframe. So I will just delete the first row.

```{r message=FALSE, warning=FALSE}

# Delete first row.
New_Dates = New_Dates[-1,]

# Look at first few rows.
head(New_Dates)

```

Now the datafrmae starts on the NEXT day.

<br/>

Next, I just added this dataframe to the modelling dataframe.

```{r message=FALSE, warning=FALSE}

# Add future dates to end of modelling dataframe.
Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)

# Look at last 15 rows of ammended dataframe.
tail(Covid_Canada_dates, n = 15)

```

Now the dataframe has contains future dates as well.

<br/>

Next comes defining the actual model. Unfortunately, I don't have any useful predictors, other than date. However, just using a date to predict the number of COVID-19 cases is probably severely lacking. Still, Using current infection trends for the past little while might be a good way to estimate very near future infection rates (like infections 5 days from now). Thus, I am going to set up a model that only uses date to predict infections for the next 10 days.

<br/>

First, a very naive model. I use a single predictor (dates), and use data from the past 300 days estimate infections.

```{r message=FALSE, warning=FALSE}

# Somewhat naive linear model.
model.1 <- lm(data = tail(Covid_Canada_dates, n = 300), cases ~ date)

```

<br/>

I can then look at a summary of the coefficients I estmated for my model.

```{r message=FALSE, warning=FALSE}

# Model Summary.
broom::tidy(model.1, conf.int = T)

```

Coefficient Interpretation: For every 1 day increase, we would expect an additional 461 new COVID-19 cases.

## Prediction
We have our model, so now let's apply those coefficients to the new days. Since our model is: COVID cases = 328270900 + date*461, we can predict cases on all new days be simply applying each date to the linear equation.

```{r message=FALSE, warning=FALSE}

# Predict COVID-19 infections with my model.
Covid_Canada_dates$New <-  predict(object = model.1, newdata = Covid_Canada_dates)

# Look at predictions for future dates.
tail(Covid_Canada_dates$New, n = 10)
```

Each value above represents predictions from our model for the next 10 days. But how accurate are these predictions really?

```{r message=FALSE, warning=FALSE}

# Plot data and future predictions regression lne.
ggplot(data = tail(Covid_Canada_dates, n = 300), aes(x = date, y = cases)) + 
  geom_point(color='blue')  +
  geom_line(color='red',data = tail(Covid_Canada_dates, n = 300), aes(x=date, y=New))

```

The regression line extends beyond the data, so it's definitely predicting future infections. But, the model isn't a very good fit as the straight line doesn't fit the non-linear data very well. A single-degree model with just one predictor doesn't allow for any flexibility. So, we could instead fit a series of polynomial models.

<br/>

In the code chunk below, I will repeat the steps I performed above, but this time fit a series of polynomial models. This will be done in a for loop, from 1 degree to 8 degrees. Finally, I will plot each model against the data.

```{r message=FALSE, warning=FALSE}

# Make 8 polynomial models, graph them, and collect adjusted R-Square value for each model.
for(i in 1:8){
  
  # Make modelling dataframe. 
  Covid_Canada_dates <- Covid_Canada
  
  # Average cases per day, for all provinces.
  Covid_Canada_dates <- Covid_Canada_dates %>%
    ddply(c("date"), summarise, 
          cases = mean(cases), 
          deaths = mean(deaths),
          DailyCases = sum(DailyCases),
          DailyDeaths = sum(DailyDeaths))
  
  # New dates dataframe.
  New_Dates <- data.frame(date =seq(max(Covid_Canada_dates$date),max(Covid_Canada_dates$date) + 30,by='days'),
                          cases = NA,
                          deaths = NA,
                          DailyCases = NA,
                          DailyDeaths = NA)
  
  # Delete first row.
  New_Dates = New_Dates[-1,]
  
  # Add future dates to end of modelling dataframe.
  Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)
  
  # Fit increasingly comlex polynomials.
  model.1 <- lm(data = tail(Covid_Canada_dates, n = 300), cases ~ poly(date, i))

  # Predict COVID-19 infections with each polynomial model.
  Covid_Canada_dates$New <-  predict(object = model.1, newdata = Covid_Canada_dates)
  
  # Assign new name to each graph I make below.
  assign(paste0("Graph", i),
         
         # Plot data and polynomial regression line.
         ggplot(data = tail(Covid_Canada_dates, n = 300), aes(x = date, y = cases)) + 
           geom_point(color='blue')  +
           ggtitle(paste0(i, " degree polynomial")) +
           theme(plot.title = element_text(size=10)) +
           theme(plot.title = element_text(hjust = 0.5)) +
           geom_line(color='red',data = tail(Covid_Canada_dates, n = 300), aes(x=date, y=New)))
  
  # Get Adjusted R-Squared value for each polynomial and assign it a new name.
  assign(paste0("Adj", i), summary(model.1)$adj.r.squared)
  
  # Model Summary.
  assign(paste0("Model_Coeffs", i),broom::tidy(model.1, conf.int = T))
  
}

```

```{r message=FALSE, warning=FALSE, fig.height=6, fig.width=7}

# Plot all 8 models on a single graph.
multi <- ggarrange(Graph1, Graph2, Graph3, Graph4, Graph5, Graph6, Graph7, Graph8)
annotate_figure(multi,top = text_grob("Predicting Next 10 Days Of COVID", color = "black", face = "bold", size = 14))

```

Each model fits differently, with the 3 order polynomial maybe looking the best in terms of future predictions. However, these models may also be severly overfitting to the data. Plus, using dates from 8 months ago might not be best when trying to assess short term trends in infection rates. So, I am going to repeat the loop above, but this this time predict the next 10 days of infection using the previous 30 days as a predictor.

```{r message=FALSE, warning=FALSE}

# Make 8 polynomial models, graph them, and collect adjusted R-Square value for each model.
for(i in 1:8){
  
  # Make modelling dataframe.
  Covid_Canada_dates <- Covid_Canada
  
  # Average cases per day, for all provinces.
  Covid_Canada_dates <- Covid_Canada_dates %>%
    ddply(c("date"), summarise,
          cases = mean(cases),
          deaths = mean(deaths),
          DailyCases = sum(DailyCases),
          DailyDeaths = sum(DailyDeaths))
  
  # New dates dataframe.
  New_Dates <- data.frame(date =seq(max(Covid_Canada_dates$date),max(Covid_Canada_dates$date) + 10,by='days'),
                          cases = NA,
                          deaths = NA,
                          DailyCases = NA,
                          DailyDeaths = NA)
  
  # Delete first row.
  New_Dates = New_Dates[-1,]

  # Add future dates to end of modelling dataframe.
  Covid_Canada_dates <- rbind(Covid_Canada_dates, New_Dates)

  # Fit increasingly comlex polynomials.
  model.2 <- lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, i))
  
  assign(paste0("model_", i), lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, i)))

  #broom::tidy(model.2, conf.int = T)

  # Predict COVID-19 infections with each polynomial model.
  Covid_Canada_dates$New <-  predict(object = model.2, newdata = Covid_Canada_dates)
  
  # Assign new name to each graph I make below.
  assign(paste0("Graph", i),
         
         # Plot data and polynomial regression line.
         ggplot(data = tail(Covid_Canada_dates, n = 30), aes(x = date, y = cases)) + 
           geom_point(color='blue')  +
           ggtitle(paste0(i, " degree polynomial")) +
           theme(plot.title = element_text(size=10)) +
            theme(axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1), 
                  plot.title = element_text(hjust = 0.5)) +
           geom_line(color='red',data = tail(Covid_Canada_dates, n = 30), aes(x=date, y=New)))
  
  # Get Adjusted R-Squared value for each polynomial and assign it a new name.
  assign(paste0("Adj", i), summary(model.2)$adj.r.squared)

  assign(paste0("Mod", i), lm(data = tail(Covid_Canada_dates, n = 30), cases ~ poly(date, i)))
  
  # Model Summary.
  assign(paste0("Model_Coeffs", i),broom::tidy(model.1, conf.int = T))

  }


```
  
```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=6}

multi <- ggarrange(Graph1, Graph2, Graph3, Graph4, Graph5, Graph6, Graph7, Graph8)
annotate_figure(multi,top = text_grob("Predicting Next 10 Days Of COVID", color = "black", face = "bold", size = 14))

```

Some of the higher order polynomial fits are pretty bad, but a few look alright. I'm going to focus on the 3-order polynomial and pretend it was the model I thought was best.

```{r message=FALSE, warning=FALSE}
# Model Summary.

broom::tidy(model_3, conf.int = T)

```

We could then apply these coefficients to each new date and get a new COVID-19 infection total for that day in the future.

<br/>

It should be stressed, however, that this model probably isn't very good. A more accurate model would need many more predictors, and I would probably assess overfitting using processes like cross-validation to assess model variance out of sample. But for this assignment, I will leave this as it is.

# Functions

The dataset I used contains data from every country in the world, not just Canada. If I wanted to repeat the graphs and analyses I did, I would have to painstakingly replace each code chunk with the new country and re-run everything. I decided to make some functions that not only repeat the analyses I did for Canada, but allow the user to simply add the country they want to the function, and instantly get up to date COVID graphs and numbers for that country.

## Function 1

### Country List
For users of these functions, it will be useful to get a list of names of countries that can be used. This function takes no arguments, and simply outputs a list of all countries the user could choose from.
```{r message=FALSE, warning=FALSE}

Country_List <- function(){
  
  # Total Global Cases Data
  x <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(x, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # compute all unique countries and return a list.
  return(unique(COVID_Global_Long_Date$`Country/Region`))

}

```

```{r message=FALSE, warning=FALSE}

# Test the function.
Country_List()

```

## Function 2

### Total cases/deaths in total country
```{r message=FALSE, warning=FALSE}

COVID_Country_Data <- function(country){
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  ### Some countries have provinces, others don't. The strategy I am using to make the dataframes for entire country
  ### won't work in both instances. So we need to apply averaging differently.
  
  # If Province/State is empty:
  if(is.na(Covid$`Province/State`) == TRUE){
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    }
  
  # If Province/State has values:
  else{
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    }
  
  # Make total cases/deaths chart.
  Chart <- Covid %>%
  ddply(c("`Country/Region`"), summarise,
        cases = sum(DailyCases),
        deaths = sum(DailyDeaths)) %>%
    gather(Category, Number, cases:deaths)
  
  # Return chart.
  return(Chart)
  
  }

```

## Test COVID Country Data Function

To make sure the function can deal both with situations where their are no provinces, and situations where their are provinces.

```{r message=FALSE, warning=FALSE}

# Test Portugal, since it has no Provinces.
COVID_Country_Data(Portugal)

# Test Canada, since it has Provinces.
COVID_Country_Data(Canada)

```

## COVID Country Graph Function

```{r message=FALSE, warning=FALSE}

COVID_Country_Graph <- function(country){
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  ### Some countries have provinces, others don't. The strategy I am using to make the dataframes for entire country
  ### won't work in both instances. So we need to apply averaging differently.
  
  # If Province/State is empty:
  if(is.na(Covid$`Province/State`) == TRUE){
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
    }
  
  # If Province/State has values:
  else{
    
    # Calculate cases per day by subtracting current count in each row by total cases up until that point.
    Covid$DailyCases <- ave(Covid$cases,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
    Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Province/State`,  FUN=function(x) c(0, diff(x)))
    
    }
  
  # Set Theme.
  ggthemr("flat")

  # average data in dataframe for graphing.
  a <- Covid %>%
    ddply(c("`Country/Region`"), summarise,
          cases = sum(DailyCases),
          deaths = sum(DailyDeaths)) %>%
    gather(Category, Number, cases:deaths) %>%
    
    # Graph total cases and deaths per province, tilt sideways, stack cases/deaths, display cases count on x-axis/death count above each bar.
    ggplot(aes(x = `Country/Region`, y = Number, fill = factor(Category, levels = c("deaths", "cases")))) +
    geom_bar(stat = "identity", position = "dodge") + 
    scale_colour_ggthemr_d() +
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE))  +
    labs(x = "Country",
         y = "COVID-19 Cases And Deaths",
         title = paste0("Total COVID-19 Cases In ", country),
         caption = paste0("Data source: Government of ", country),
         fill = "Case Type") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label = Number), position = position_dodge(width=0.9), vjust = -.2)
  
  return(a)
  
}

```

## Test COVID Country Graph Function

To make sure the function can deal both with situations where their are no provinces, and situations where their are provinces.

```{r message=FALSE, warning=FALSE}

# Test Portugal, since it has no Provinces.
COVID_Country_Graph(Portugal)

# Test Canada, since it has Provinces.
COVID_Country_Graph(Canada)

```

## COVID Country Cumulative Graph Function

```{r message=FALSE, warning=FALSE}

COVID_Country_Cumulative_Graph <- function(country){
  
  # Adds quotes to country argument which is needed for the subset function.
  country <- rlang::quo_name(rlang::enquo(country))
  
  # Total Global Cases Data
  COVID_Global_Wide<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

  # Total Global Deaths Data
  COVID_Global_Deaths_Wide <-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))
  
  # Total global cases data in long format.
  COVID_Global_Long_Date <- gather(COVID_Global_Wide, date, cases, 5:ncol(COVID_Global_Wide), factor_key=TRUE)
  
  # Total global deaths data in long format
  COVID_Global_Deaths_Long_Date <- gather(COVID_Global_Deaths_Wide, date, cases, 5:ncol(COVID_Global_Deaths_Wide), factor_key=TRUE)
  
  # Combine death count to cases dataframe.
  COVID_Global_Long_Date$deaths <- COVID_Global_Deaths_Long_Date$cases
  
  # Filter canadian provincial data to only include data starting from March 1st.
  Covid <- Covid[352:nrow(COVID_Global_Long_Date), ]
  
  # Subset Canadian provinces from total cumulative global cases data.
  Covid <- subset(COVID_Global_Long_Date, `Country/Region` == country)
  
  # Convert date column to date format for cases dataframe.
  Covid$date <- as.Date(Covid$date, format = "%m/%d/%Y")
  
  # Calculate cases per day by subtracting current count in each row by total cases up until that point.
  Covid$DailyCases <- ave(Covid$cases,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
    
  # Calculate deaths per day by subtracting current count in each row by total deaths up until that point.
  Covid$DailyDeaths <- ave(Covid$deaths,Covid$`Country/Region`,  FUN=function(x) c(0, diff(x)))
  
  # Set Theme.
  ggthemr("flat")
  
  # Cumulative Cases Graph.
  a <- Covid %>% 
    ggplot(aes(x = date, y = cases, color = `Country/Region`, group = `Country/Region`, fill = `Country/Region`)) +
    geom_area() +
    labs(x = "March 1st (2020) To Present",
         y = "Cumulative Cases Per 100,000 People",
         title = "Coronavirus Cases Per Canadian Province",
         caption = "Data source: Government of Canada") +
    scale_colour_ggthemr_d() +
    scale_x_date(date_breaks = '1 month', date_labels = '%B') +
    theme(axis.text.x = element_text(face = "bold", size = 9, angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))
  
  return(a)
  
}

```

```{r message=FALSE, warning=FALSE}

#COVID_Country_Cumulative_Graph(Portugal)
#COVID_Country_Cumulative_Graph(US)

```